---
title: "Determining What Factors Make A Product Expensive Or Not"
author: "Fang Yang, Jonathan Whittle, Kainin Liu, Ziqi Yang and Ziyao Zong"
date: "2023-03-08"
output:
  pdf_document:
    number_sections: yes
  html_document:
    number_sections: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(tinytex.verbose = TRUE)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#reading data
data <- read.csv("dataset22.csv")


#loading libraries
library(tidyverse)
library(moderndive)
library(sjPlot)
library(stats)
library(jtools)
library(GGally)
library(janitor)
library(gridExtra)
library(kableExtra)
library(skimr)
library(mice)
library(Hmisc) 
library(dplyr)
library(glm2)
```

# Introduction {#sec:intro}

We have been tasked with finding out which properties of the furniture influence whether the products price is more than 1000 Saudi Riyals. Here we shall analyse the data provided to us by the company, working with continuous variables for price, width, depth and height as well as categorical variables for category, whether the product is sold online and whether the product is available in multiple colours.

Section \ref{sec:eda} consists of an exploratory data analysis of the data. Section \ref{sec:fda} contains the contains the formal analysis of fitting the General linear model using logistic regression, the model selection and data visualizations resulting from the model selected. Section \ref{sec:con} contains our conclusions. Before the Exploratory Data Analysis part we first address the missing values. From the table \ref{tab:summaryna} we can see that our "depth" data have 213 missing data points which is almost 43 percent of the dataset so we first delete the rows where all of depth, height and width are missing, leaving us with 443 observations. We then replaced the missing data using the Multiple Imputation method.

```{r, echo=FALSE, warning=FALSE, message=FALSE,results='markup', fig.show='asis', eval=TRUE, table.pos="H"}

data_na<-data %>% select(depth, height, width)

summary_data <- data_na %>%
  summarise_all(funs(na_count = sum(is.na(.))))

summary_table <- summary_data %>%
  kable(caption = '\\label{tab:summaryna} Summary table of the IKEA Saudi Arabia missing data') %>%
  kable_styling(font_size = 10, latex_options = "hold_position")

summary_table

```

```{r results = 'hide', echo=FALSE, warning=FALSE, message=FALSE}
#data tidying

#delete the row which both depth, height and width are NA
nas <- c()
coun = 1
for(i in 1:nrow(data)){
  if(is.na(data$depth[i])){
    if(is.na(data$height[i])){
      if(is.na(data$width[i])){
        nas[coun] = i
        coun = coun + 1
      }
    }
    else
      next
  }
  else
    next
}
data <- data[-nas, ]
data = data[, -c(1:2)]
#splitting price into categories expensive yes or no
data$Expensive <- NA
data$Expensive[data$price<1000] <- 0
data$Expensive[data$price>=1000] <- 1
data$Expensive <- factor(data$Expensive, labels=c("No", "Yes"))
#use Multiple Imputation to replace NA values
set.seed(1)
imp <- mice(data, m=5, maxit = 50)
#change imp to dataframe
data <- complete(imp)

#tidying category with irregular symbol
#unique(data$category)
#data[131,1]
data[131,1] = "Cafe furniture"

```

# Exploratory Data Analysis {#sec:eda}

As we are interested in whether a product is sold at more or less than 1000 Riyals, we dichotamised the variable for price into a categorical variable: Expensive. Products sold for more than 1000 Riyals are denoted "Yes" and products sold for less than 1000 Riyals are denoted "No". Below we provide a summary table \ref{tab:summariesskim} for the numerical data and a table \ref{tab:percentage} showing the proportion of products considered expensive. From the summary table below we can see that the mean price is over 1000 Riyals.

```{r, echo=FALSE, warning=FALSE, message=FALSE,results='markup', fig.show='asis', eval=TRUE, table.pos="H"}
#create ikea_num data for numerical data
ikea_num<-data %>% select(price, depth, height, width)
#summary table
my_skim <- skim_with(numeric = sfl(hist = NULL), 
                    base = sfl(n = length))
my_skim(ikea_num) %>%
  transmute(Variable=skim_variable, n = n, Mean=round(numeric.mean,2), SD=round(numeric.sd,2),
            Min=numeric.p0, Median=numeric.p50,  Max=numeric.p100,
            IQR = numeric.p75-numeric.p50) %>%
  kable(caption = '\\label{tab:summariesskim} Summary statistics 
        of the IKEA Saudi Arabia data') %>%
  kable_styling(font_size = 10, latex_options = "hold_position")

#table for overall expensive
tabyl(data$Expensive) %>%
  kable(caption= '\\label{tab:percentage} Counts and percentages of products sorted by whether they are expensive or not') %>%
  kable_styling(font_size = 10, latex_options = "hold_position")

```

We now turn our attention to which variables influence whether a product is considered expensive or not. From the boxplots, Figure \ref{fig:boxplot1}, Figure \ref{fig:boxplot2}, Figure \ref{fig:boxplot3} below we can see that in terms of the continuous variables width appears to be the most important.

```{r, echo=FALSE, warning=FALSE, message=FALSE,out.width="60%",fig.align = "center",fig.pos = "H",fig.cap="\\label{fig:boxplot1} Boxplot of expensive and width",fig.show='asis'}
#plots

#boxplot of width and expensive
p1 <- ggplot(data= data, aes(x=width, y = Expensive, na.rm=TRUE)) +
  geom_boxplot(fill ="steelblue") +
  labs(x="Width", y="Expensive")
p1
```

```{r, echo=FALSE, warning=FALSE, message=FALSE,out.width="60%",fig.align = "center",fig.pos = "H",fig.cap="\\label{fig:boxplot2} Boxplot of expensive and depth",fig.show='asis'}
#boxplot of depth and expensive
p2 <- ggplot(data= data, aes(x=depth, y = Expensive, na.rm=TRUE)) +
  geom_boxplot(fill = "steelblue") +
  labs(x="Depth", y="Expensive")
p2
```

```{r, echo=FALSE, warning=FALSE, message=FALSE,out.width="60%",fig.align = "center",fig.pos = "H",fig.cap="\\label{fig:boxplot3} Boxplot of expensive and height",fig.show='asis'}
#boxplot of height and expensive
p3 <-ggplot(data= data, aes(x=height, y = Expensive, na.rm=TRUE)) +
  geom_boxplot(fill= "steelblue") +
  labs(x="Height", y="Expensive")
p3
```

Checking the continuous variables for any high collinearity, we see from the scatterplot matrix, Figure \ref{fig:cor} that there is some weak positive correlation between depth and width and also between depth and height. As the strength is weak and the correlation between height and depth is very weak, there is not much concern regarding collinearity.

```{r, echo=FALSE, warning=FALSE, message=FALSE,eval = TRUE,out.width="70%",fig.align = "center",fig.pos = "H",fig.cap="\\label{fig:cor} Scatterplot matrix",fig.show='asis'}

#producing ggpairs to check for high collinearity
continuousdata <- data %>%
  select(width, depth, height)

ggpairs(continuousdata, title="Correlation of continous variables") 

```

We created boxplots looking at price for the categorical variables Sold Online and Multiple colours. From these Figure \ref{fig:Price} we can see that furniture being sold in multiple colours makes little impact on price while there is some suggestion that products sold online are more expensive.


```{r, echo=FALSE, warning=FALSE, message=FALSE,eval = TRUE,out.width="70%",fig.align = "center",fig.pos = "H",fig.cap="\\label{fig:Price} Boxplot of price and categorical variables",fig.show='asis'}
#boxplot 
b1<-ggplot(data = data, aes(x = other_colors, y = price, fill = other_colors)) +
  geom_boxplot() +
  labs(x = "Other colours", y = "Price")+
  theme(legend.position="bottom")


b2<-ggplot(data = data, aes(x = sellable_online, y = price, fill = sellable_online)) +
  geom_boxplot() +
  labs(x = "Sellable online", y = "Price")+
  theme(legend.position="bottom")

grid.arrange(b1,b2,ncol=2)
```

From further research of categorical variables, we can see from table \ref{tab:sellable} that as so few products are not sold online, it is unlikely to be influential. And from the barplot, Figure \ref{bar:categorical} as the proportions for expensive are very similar for multiple colours, it is also unlikely to have much influence.

```{r, echo=FALSE, warning=FALSE, message=FALSE, eval = TRUE,out.width="70%",fig.align = "center",fig.pos = "H",fig.cap="\\label{bar:categorical}  Barplot of categorical variables",fig.show='asis'}

#categorical variables

#expensive and multiple colours
#table 
t1 <- data %>% 
  tabyl(Expensive, other_colors) %>% 
  adorn_percentages() %>% 
  adorn_pct_formatting() %>% 
  adorn_ns() %>%
  kable(caption='\\label{tab:colours} Percentages and counts of products separated whether they have multiple colours ') %>%
  kable_styling(font_size = 10, latex_options = "hold_position")
t1
#barplot
bp1 <- ggplot(data, aes(x = other_colors, fill = Expensive, na.rm=TRUE)) +
  geom_bar(stat = "count", position = "dodge") +
  xlab("Other colours") + ylab("Frequency") +
  theme(legend.position="bottom")
#expensive and sold online
#table
t2 <- data %>% 
  tabyl(Expensive, sellable_online) %>% 
  adorn_percentages() %>% 
  adorn_pct_formatting() %>% 
  adorn_ns() %>%
  kable(caption= '\\label{tab:sellable} Percentages and counts of products separated by whether they are sold online') %>%
  kable_styling(font_size = 10, latex_options = "hold_position")
t2
#barplot
bp2 <-ggplot(data, aes(x = sellable_online, fill = Expensive, na.rm=TRUE)) +
  geom_bar(stat = "count", position = "dodge")+
  xlab("Sellable Online") + ylab("Frequency") +
  theme(legend.position= "bottom")
grid.arrange(bp1,bp2,ncol=2)
```

From the plot, Figure \ref{bar:category} we see that categories is likely to be an influential factor. Wardrobes, Sofas and Armchairs and beds in particular have a high proportion of expensive products. Meanwhile Childrens furniture contains no expensive products at all. Some categories contain so few products that they are unlikely to be influential.

```{r, echo=FALSE, warning=FALSE, message=FALSE, eval = TRUE,out.width="100%",fig.align = "center",fig.pos = "H",fig.cap="\\label{bar:category}  Barplot of expensive and not expensive categories",fig.show='asis'}
#expensive and categories
#table
t3 <- data %>% 
  tabyl(category, Expensive) %>% 
  adorn_percentages() %>% 
  adorn_pct_formatting() %>% 
  adorn_ns() %>%
  kable(caption= '\\label{tab:category} Percentages and counts of products sorted by category') %>%
  kable_styling(font_size = 10, latex_options = "hold_position")
t3
#barplot
ggplot(data, aes(y=category,  fill = Expensive, na.rm=TRUE)) +
  geom_bar(stat = "count", position = "dodge", width=0.5) +
  ylab("Category") + xlab("Frequency") +
  theme(legend.position = "bottom") +
  ggtitle("Counts of expensive and not expensive categories")
```
\newpage

# Formal analysis {#sec:fda}
To model are data we fitted a GLM using logistic regression using the logit link function. After starting with a model containing the main effects for the variables height, width, depth, categories, sellable_online and other_colors. We found that sellable_online was the least significant with a p-value of 1.00. The BIC for this model was and 382.7 respectively. We then fitted a model without the term for sellable_online. This second model had a lower value for BIC of 377.3 so we preferred this second model to the first. However, categories was not significant as all p-values were 1.00. We then fitted a third model with variables height, depth, width and other_colors. The BIC for this third model was 333.4 so the third model is preferred. The variable other_colors was not significant with a p-value of 0.19, so we fitted a fourth model with variables height, width and depth. This model had a BIC value of 329.1 so was preferred to the third model. Three models were then fitted with interaction terms for height and depth, depth and width and finally width and height respectively. However, none of these terms were significant in their models and all three models resulted in an increase in BIC so the model with variables for the main effects of height, width and depth was selected.


```{r,echo=FALSE, warning=FALSE, message=FALSE, fig.align = "center",fig.pos = "H",fig.show='asis' }

#model full
data1<-data %>% select(category, sellable_online, other_colors, depth, height, width, Expensive)
model_full <- glm2(Expensive~., data = data1, family = binomial(link = "logit"))
#summ(model_full)
#AIC 292.66,BIC 382.72, least significant term Sellable_online with p-value 1.00

#model without sellable_online
data2<- data %>% select(category, other_colors, depth, height, width, Expensive)
model2 <- glm2(Expensive~., data = data2, family = binomial(link = "logit"))
#summ(model2)
#AIC 291.37, BIC 377.34, least significant term category with p-values 0.99

#model without categories
data3 <- data %>% select(other_colors, height, width, depth, Expensive)
model3 <- glm2(Expensive~., data = data3, family = binomial(link = "logit"))
#summ(model3)
#AIC 312.97, BIC 333.44, least significant term other_colors with p-value 0.19

#model without other_colors
data4 <- data %>% select(height, width, depth, Expensive)
model4 <- glm2(Expensive~., data = data4, family = binomial(link = "logit"))
summ(model4)
#AIC  312.71, BIC 329.08, all terms significant

#trying models with an interaction
mod5 <- glm2(Expensive~ height*depth + width,data = data4, family = binomial(link = "logit"))
#summ(mod5)
#AIC 313.17, BIC 333.64
mod6 <- glm2(Expensive~ depth*width + height,data = data4, family = binomial(link = "logit"))
#summ(mod6)
#AIC 310.30, BIC 330.77
mod7 <- glm2(Expensive~ width*height + depth,data = data4, family = binomial(link = "logit"))
#summ(mod7)
#AIC 314.40, BIC 334.86
#none of these models are an improvement so we use model4
```


Thus the best fitted model is as follows:
$$
\ln(\displaystyle\frac{p}{1-p})=\alpha+\beta_{1}\cdot\textrm{height}+ \beta_{2} \cdot \textrm{width}+ \beta_{3} \cdot \textrm{depth}
$$
$$
=-6.65+0.01\cdot\textrm{height}+ 0.02\cdot \textrm{width}+ 0.05 \cdot \textrm{depth}
$$

where $p$ = Prob(price over or equal to 1000) and 1 - $p$ = Prob(price under 1000)
We see that the intercept is negative while the coefficients for height, width and depth are positive. Hence, the log-odds of the price being expensive increase by 0.01 for every one unit increase in height, 0.02 for width, 0.05 for depth. This suggests that an increase in depth leads to the highest increase in probability that a product is expensive.


```{r, echo=FALSE, fig.align='center', warning=FALSE, message=FALSE, results='hide'}
#obtaining coefficients
modcoefs <- round(coef(model4), 2)
```

\newpage
From the table \ref{tab:con} we can see that all the intervals do not contain zero so all variables are significant.

```{r, r, echo=FALSE, warning=FALSE, message=FALSE,results='markup', fig.show='asis', eval=TRUE, table.pos="H"}
confint(model4) %>%
  kable(caption = '\\label{tab:con} Confidence interval table') %>%
  kable_styling(font_size = 10, latex_options = "hold_position")
```



```{r, echo=FALSE, warning=FALSE, message=FALSE,eval = TRUE,out.width="70%",fig.align = "center",fig.pos = "H",fig.cap="\\label{fig:logodds} Log-Odds of expensive",fig.show='asis'}
#plotting log odds
plot_model(model4, show.values = TRUE, transform = NULL, show.p = FALSE)

#adding log odds into data
data <- data %>%
  mutate(logodds.expensive = predict(model4))
```

Figure \ref{fig:logodds}, a plot of the log-odds for height, width and depth.

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='hide'}
#creating odds
model4 %>%
  coef() %>%
  exp()
```

The odds of the model is as follows
$$
\displaystyle\frac{p}{1-p}=exp(\alpha+\beta_{1}\cdot\textrm{height}+ \beta_{2} \cdot \textrm{width}+ \beta_{3} \cdot \textrm{depth})\
$$
$$
=0.0013+1.01\cdot\textrm{height}+ 1.02\cdot \textrm{width}+1.05\cdot \textrm{depth}
$$
This suggests that the coefficients for height, width and depth are all positive. And for each increase of 1 unit in height the odds of being expensive are 1.01 times those
of the same height and for width is 1.02 times, depth is 1.05 times.

A plot of the odds for height, width and depth is shown below, Figure \ref{fig:odds}.

```{r, echo=FALSE, warning=FALSE, message=FALSE,eval = TRUE,out.width="70%",fig.align = "center",fig.pos = "H",fig.cap="\\label{fig:odds} Odds in expensive",fig.show='asis'}


#plotting odds
plot_model(model4, show.values = TRUE, axis.lim = c(1,1.5), show.p = FALSE)

#adding odds and probabilities to data
data <- data %>%
  mutate(odds.expensive = exp(logodds.expensive)) %>%
  mutate(prob.expensive = fitted(model4))
```
below plots showing how the probability of a product being expensive varies with the explanatory variables: height, depth and width.

```{r, echo=FALSE, warning=FALSE, message=FALSE, eval = TRUE,out.width="70%",fig.align = "center",fig.pos = "H",fig.cap="\\label{fig:probplot} Probability plots",fig.show='asis'}
#plotting probabilities
#height
prob1 <- ggplot(data = data, aes(x = height, y = prob.expensive)) +
  geom_smooth(method="glm", 
              method.args = list(family="binomial"), 
              se = FALSE) +
  labs(x = "height", y = "P(expensive)")
#depth
prob2 <- ggplot(data = data, aes(x = depth, y = prob.expensive)) +
  geom_smooth(method="glm", 
              method.args = list(family="binomial"), 
              se = FALSE) +
  labs(x = "depth", y = "P(expensive)")
#width
prob3 <- ggplot(data = data, aes(x = width, y = prob.expensive)) +
  geom_smooth(method="glm", 
              method.args = list(family="binomial"), 
              se = FALSE) +
  labs(x = "width", y = "P(expensive)")

grid.arrange(prob1,prob2,prob3, ncol=1)
```
\newpage
# Conclusions {#sec:con}

Finally we can answer the question: Which properties of furniture influence whether they cost more than 1000 Saudi Riyals? 
From the model we can see that, the furniture's category, whether they sell online and whether it have different colors may not have significant influence on the price over 1000 Saudi Riyals. While the height, depth and width of the furniture all have significant influence on the whether the price is over 1000 Saudi Riyals, the furniture's depth has the greatest influence. As the furniture become larger (the values for height, depth and width increase) the probability that its price will be more than 1000 Saudi Riyals will increase, eventually converging to one.


